---
title: "Assignment 1"
author: "Vera Weidmann, Marvin König, Sebastian Seck"
date: "5/22/2017"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("~/Documents/C Education/3 HWR/2. Semester SS2017/ECO")
```

```{r Initialization, warning=FALSE}
library(foreign) #to read in the .dta format
library(stargazer) #used in 1(c)-ii
hprice <- read.dta("/Users/marvin/Documents/Hochschule für Wirtschaft & Recht/Business Intelligence & Process Management/Econometrics/Data/hprice1.dta")
wage2 <- read.dta("/Users/marvin/Documents/Hochschule für Wirtschaft & Recht/Business Intelligence & Process Management/Econometrics/Data/WAGE2.DTA")
```

##Task 1
#### 1(a)
```{r 1a}
names(hprice)
summary(hprice[,c("price","bdrms","lotsize")])
```
The sample mean of price is 293.5. The summary statistics can be found above.

#### 1(b)
```{r 1b}
mod1b <- lm(lprice ~ bdrms + llotsize, data=hprice)
summary(mod1b)
```
The summary provides us with an R-Squared of 40.13% - this means that 40% of deviation in (log-) price can be explained by the predictors chosen. We are also provided with an adjusted R-squared, which is adjusted not to consider the count of variables used for the model.
n can be derived from the degrees of freedom: $DF = n-k-1 \Leftrightarrow 85 = 88-2-1$ From this formula we can easily deduct the number of observations n, which is 88.

If we consider that bdrms and lotsize is equal to 0, the price is the value of the intercept: 2.95. The model isa log-level-model. This means, an increase of one predictor by holding the others fixed leads to an percentage increase of the response varible (log-price).
So, if bdrms is increasing by one extra room while the other predictor lotsize is held fixed, the price increases by 14%. For one increase in log(lotsize) the price rise by 24.4%.

$\frac{\delta log(price)}{\delta bdrms}=\beta_1$
$\frac{\delta log(price)}{\delta log(lotsize)}=\beta_2$

Furthermore, both coefficients are significant. The p-vales of both coefficients are one indicator for the significance. As they are lower than $\alpha = 5%$ (and $\alpha = 1%$), we can reject that both predictors has no effect on the response price (H0: $\beta_1 = 0, \beta_2  =0$) in a two-sided test.
Likewise we could look at the t-values for signifcance. In R the significance codes give also a clarification about the coef. significance.

#### 1(c) - i
```{r 1ci}
#1(c) - i
hprice$sqllotsize <- hprice$llotsize^2 #to account for the additional variable, we simply create another column
mod1c <- lm(lprice ~ bdrms  + llotsize + sqllotsize, data = hprice)
summary(mod1c)
```
#### 1(c) - ii
```{r 1cii, results="asis"}
stargazer(mod1b,mod1c, title = "Comparison of models mod1b and mod1c", type="html", model.names = FALSE,column.labels = c("mod1b", "mod1c"), column.separate = c(1,1), style = "qje")
```

The first derivative of $log(price)=\beta_0+\beta_1bdrms+\beta_2log(lotsize)+\beta_3[log(lotsize)]^2+u$:  

$\frac{\delta log(price)}{\delta log(lotsize)}=\beta_2+2\beta_3log(lotsize)$

In model "mod1c" we add the squared value of llotsize. As it is shown in the derivation above, the effect from lotsize by one unit depends on how many units of lotsize is already there. So if... EXAMPLE!!!

#### 1(e)
At first we look at the variance of the error term of model 2, for which we square the Root MSE. In order to calculate the standard error of llotsize and bdrms we are using the SST and R-squared from the respective model outputs, where the variables are the dependent variable. We take the square root in order to arrive at the standard error instead of the variance.

In detail, this is how $SE_{\beta_2}$ (llotsize) is calculated:  
$SE_{\beta_2}=\sqrt{\frac{\sigma^2}{SST_2(1-R^2_2)}} = \sqrt{\frac{0.23^2}{25.75*(1-0.9939)}}=0.604$

In detail, this is how $SE_{\beta_1}$ (bdrms) is calculated:  
$SE_{\beta_1}=\sqrt{\frac{\sigma^2}{SST_1(1-R^2_1)}} = \sqrt{\frac{0.23^2}{61.59*(1-0.0289)}}=0.0308$

##Task 2
####2(a)
```{r 2a}
mod2a <- lm(lwage ~ educ + exper + tenure + married + black + south + urban, data = wage2)
summary(mod2a)

wage2$nonblack <- 1 - wage2$black

#90% Confidence Interval (alpha = 0.05 two-sided)
x = mod2a$coefficients["black"]
SE = sqrt(diag(vcov(mod2a)))["black"]
CI1= x - SE * 1.64
CI2 = x + SE * 1.64
(salary_black <-cbind(CI1, CI2))
```
The salary difference can be found from the coefficient $\beta_5$ in model mod2a, it is -0.1884. As the CI is not included in R's standard model output, we calculated it manually.

####2(b)
```{r 2b}
#Creating the additional variables
wage2$exper_sq <- wage2$exper^2
wage2$tenure_sq <- wage2$tenure^2

mod2b <- lm(lwage ~ educ + exper + tenure + married + black + south + urban + exper_sq + tenure_sq, data = wage2)
summary(mod2b)

#Calculating the residual sum of squares for both models
SSRr <- sum(mod2a$residuals^2)
SSRur <- sum(mod2b$residuals^2)

SSRr
SSRur
```

Our null hypothesis is that both additional variables have no explanatory power. So, model 2a (the smaller one) would be better than model 2b. 

$H_0: \beta_8=0; \beta_9=0$
$H_1: \beta_8>0; \beta_9>0$

Dicision Rule:
Reject H0 if $F > 1.89$ ($\alpha = 0.05$, one sided test) 

In R we also can calculate the critical value instead of using the table of the F-distributions:
```{r}
#(1-/alpha, k, n-k-1)
qf(0.95, 9, 925)
```

To calculate the F-statistic, we use the following formula:
$F = \frac{(SSR_r - SSR_{ur}) / q}{SSR_{ur} / (n - k - 1)}$

In this case the restricted model is 2a and the unrestricted one is model 2b. 

```{r}
q = 2
n = 935 
k= 9

(F_test=((SSRr-SSRur) / q)/(SSRur / (n-k-1)))
```

Substituted with our values:
$F = \frac{(123.8185 - 123.421) / 2}{123.421 / (935 - 9 - 1)} = 1.489806$

We cannot reject our H0, because F < 1.89 and lies inside the 95% Interval. This means $\beta8$ and $\beta9$ do not provide additional explanation power.
This is also seen in the output of model 2b. The p-values of these both coefficients are greater than $\alpha = 0.05$. So, exper_sq and tenure_sq might not be signifcant for our model.

-> beta9 ist gerade noch so significant bei onesided test 5% = 0.09/2

Furthermore, we can calcualte the Anova of these both models. The result underlines the previous expanation of the non-rejection of H0.

```{r}
anova(mod2a,mod2b)
```

####2(c)
```{r 2c}
mod2c <- lm(lwage ~ educ*black + exper + tenure + married + south + urban, data = wage2)
summary(mod2c)
```

Our null hypothesis is that the interaction term between education and black has no effect on the model response. 

$H0: \beta8 =0$
$H1: \beta8 != 0$ # sebastian, please fix unequal sign!!! :) 

Decision Rule: Reject H0 if |t| > 1.64   ($\alpha = 0.1 $, two sided test)

As t(educ:black) is |-1.121| < 1.64 (as well as: p-value of 26% > 10%), we cannot reject H0. The interaction term between education and black might not have an effect on the model. So, education might not depend on race.

#### 2(c) - i
```{r 2ci, results="asis"}
stargazer(mod2a,mod2c, title = "Comparison of models mod2a and mod2c", type="html", model.names = FALSE,column.labels = c("mod2a", "mod2c"), column.separate = c(1,1), style = "qje")
```

In model 2c we can see that the black coefficient is positive now. But this variable interacts with eduaction. This means being black causes an effect of +0.09-0.023*educ. So the wage of a black person depends in this case of his education. Model 2a do not integrate the interaction term. Therefore, being black decrease the persons wage by 19%.

$\frac{\delta log(wage)}{\delta black}=\beta_2+\beta_1educ$

However, the standard error for black in model 2c is quite big and we can't verify a significance for this coefficient anymore. As already seen in the hypothesis test above ($H0: \beta8 = 0 $), education also does not depend on the variable black and the model 2a might be the one which reflects the population in a better way. 

#### 2(c) - ii
```{r 2cii}
c <- mod2c$coefficients["educ"]
educ_c <- wage2$educ - c

mod2cii <- lm(lwage ~ educ_c*black + exper + tenure + married + south + urban, data = wage2)
summary(mod2cii)
```

The value we suggest is equal to estimate of education. Interpreting models with interactions can be tricky due to the fact that $\beta2$ (coefficient of black) just shows the partical effect of being black when his eduaction = 0. While using the equation black*(educ-c) we centring the variables around the mean value of education. 

#### 2(d)
```{r 2d}
mod2d <- lm(lwage ~ educ + black*married + exper + tenure + south + urban, data = wage2)
summary(mod2d)
````

Being not married and not black is our base model, which is illustrated by the intercept. 
The second group (not_married & black) is represented by the coefficient $\beta2 = -0.240820$
The coefficient $\beta3 = 0.188915$ presents the third group: being married and not being black. 
Being married and black is represented by $\beta2 + \beta3 + \beta8$. 

Therefore, the difference between the second and the forth group are  $\beta2 + \beta8 = -0.24 + 0.06 = -0.18$. (Same as in mod2a)
