---
title: "Assignment 1"
author: "Vera Weidmann, Marvin König, Sebastian Seck"
date: "5/22/2017"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("~/Documents/C Education/3 HWR/2. Semester SS2017/ECO")
```

```{r Initialization, warning=FALSE, message=FALSE}
library(foreign) #to read in the .dta format
library(stargazer) #used in 1(c)-ii
hprice <- read.dta("hprice1.dta")
wage2 <- read.dta("WAGE2.DTA")
ksubs <- read.dta("401ksubs.dta")
```

##Task 1
```{r}
head(hprice)
```

#### 1(a)
```{r 1a}
names(hprice)
summary(hprice[,c("price","bdrms","lotsize")])
```
The sample mean of price is 293.5. The summary statistics can be found above.

#### 1(b)
```{r 1b}
mod1b <- lm(lprice ~ bdrms + llotsize, data=hprice)
summary(mod1b)
```
The summary provides us with an R-Squared of 40.13% - this means that 40% of deviation in (log-) price can be explained by the predictors chosen. We are also provided with an adjusted R-squared, which is adjusted not to consider the count of variables used for the model.
n can be derived from the degrees of freedom: $DF = n-k-1 \Leftrightarrow 85 = 88-2-1$ From this formula we can easily deduct the number of observations n, which is 88.

If we consider that bdrms and lotsize is equal to 0, the price is the value of the intercept: 2.95. The model isa log-level-model. This means, an increase of one predictor by holding the others fixed leads to an percentage increase of the response varible (log-price).
So, if bdrms is increasing by one extra room while the other predictor lotsize is held fixed, the price increases by 14%. For one increase in log(lotsize) the price rise by 24.4%.

$\frac{\delta log(price)}{\delta bdrms}=\beta_1$
$\frac{\delta log(price)}{\delta log(lotsize)}=\beta_2$

Furthermore, both coefficients are significant. The p-vales of both coefficients are one indicator for the significance. As they are lower than $\alpha = 5%$ (and $\alpha = 1%$), we can reject that both predictors has no effect on the response price (H0: $\beta_1 = 0, \beta_2  =0$) in a two-sided test.
Likewise we could look at the t-values for signifcance. In R the significance codes give also a clarification about the coef. significance.

#### 1(c) - i
```{r 1ci}
#1(c) - i
hprice$sqllotsize <- hprice$llotsize^2 #to account for the additional variable, we simply create another column
mod1c <- lm(lprice ~ bdrms  + llotsize + sqllotsize, data = hprice)
summary(mod1c)
```
#### 1(c) - ii
```{r 1cii, results="asis"}
stargazer(mod1b,mod1c, title = "Comparison of models mod1b and mod1c", type="html", model.names = FALSE,column.labels = c("mod1b", "mod1c"), column.separate = c(1,1), style = "qje")
```

The first derivative of $log(price)=\beta_0+\beta_1bdrms+\beta_2log(lotsize)+\beta_3[log(lotsize)]^2+u$:  

$\frac{\delta log(price)}{\delta log(lotsize)}=\beta_2+2\beta_3log(lotsize)$

In model "mod1c" we add the squared value of llotsize. As it is shown in the derivation above, the effect from lotsize by one unit depends on how many units of lotsize is already there. So if... EXAMPLE!!!

#### 1(d)
The unbiased estimator usually implemented is $\frac{SSR}{n-k-1}$.
Generally, unbiasedness means that there is an average trend within repeated samples. Unbiasedness in this case means that the restrictions given by the OLS first order conditions are respected - which means an adjustment to the degrees of freedom.


The two suggested variance estimates for mod1c are: 
```{r 1d}
(sum(mod1c$residuals^2)/(nrow(hprice)-1))
(sum(mod1c$residuals^2)/(nrow(hprice)-4))
```
There are no large differences because the bias of the first variance estimate is diminished by a rather large n.

In order to compare this value with out model output, we take the square root of the calculated error variance to compare it with the Residual Standard Error output in our model - which are printed out for you in the following:
```{r 1d-2}
sqrt(sum(mod1c$residuals^2)/(nrow(hprice)-1))
sqrt(sum(mod1c$residuals^2)/(nrow(hprice)-4))
```


#### 1(e)
At first we look at the variance of the error term of model 2, for which we square the Root MSE. In order to calculate the standard error of llotsize and bdrms we are using the SST and R-squared from the respective model outputs, where the variables are the dependent variable. We take the square root in order to arrive at the standard error instead of the variance.

In detail, this is how $SE_{\beta_2}$ (llotsize) is calculated:  
$SE_{\beta_2}=\sqrt{\frac{\sigma^2}{SST_2(1-R^2_2)}} = \sqrt{\frac{0.23^2}{25.75*(1-0.9939)}}=0.604$

In detail, this is how $SE_{\beta_1}$ (bdrms) is calculated:  
$SE_{\beta_1}=\sqrt{\frac{\sigma^2}{SST_1(1-R^2_1)}} = \sqrt{\frac{0.23^2}{61.59*(1-0.0289)}}=0.0308$

##Task 2
```{r}
head(wage2)
```

####2(a)
```{r 2a}
mod2a <- lm(lwage ~ educ + exper + tenure + married + black + south + urban, data = wage2)
summary(mod2a)

wage2$nonblack <- 1 - wage2$black

#90% Confidence Interval (alpha = 0.05 two-sided)
x = mod2a$coefficients["black"]
SE = sqrt(diag(vcov(mod2a)))["black"]
CI1= x - SE * 1.64
CI2 = x + SE * 1.64
(salary_black <-cbind(CI1, CI2))
```
The salary difference can be found from the coefficient $\beta_5$ in model mod2a, it is -0.1884. As the CI is not included in R's standard model output, we calculated it manually.

####2(b)
```{r 2b}
#Creating the additional variables
wage2$exper_sq <- wage2$exper^2
wage2$tenure_sq <- wage2$tenure^2

mod2b <- lm(lwage ~ educ + exper + tenure + married + black + south + urban + exper_sq + tenure_sq, data = wage2)
summary(mod2b)

#Calculating the residual sum of squares for both models
SSRr <- sum(mod2a$residuals^2)
SSRur <- sum(mod2b$residuals^2)

SSRr
SSRur
```

Our null hypothesis is that both additional variables have no explanatory power. So, model 2a (the smaller one) would be better than model 2b. 

$H_0: \beta_8=0; \beta_9=0$
$H_1: \beta_8>0; \beta_9>0$

Decision Rule:
Reject H0 if $F > 1.89$ ($\alpha = 0.05$, one sided test) 

In R we also can calculate the critical value instead of using the table of the F-distributions:
```{r}
#(1-/alpha, k, n-k-1)
qf(0.95, 9, 925)
```

To calculate the F-statistic, we use the following formula:
$F = \frac{(SSR_r - SSR_{ur}) / q}{SSR_{ur} / (n - k - 1)}$

In this case the restricted model is 2a and the unrestricted one is model 2b. 

```{r}
q = 2
n = 935 
k= 9

(F_test=((SSRr-SSRur) / q)/(SSRur / (n-k-1)))
```

Substituted with our values:
$F = \frac{(123.8185 - 123.421) / 2}{123.421 / (935 - 9 - 1)} = 1.489806$

We cannot reject our H0, because $F < 1.89$ and lies inside the 95% Interval. This means $\beta_8$ and $\beta_9$ do not provide additional explanation power.

Furthermore, we can calcualte the Anova of these both models. The result underlines the previous explanation of the non-rejection of $H_0$.

```{r}
anova(mod2a,mod2b)
```

####2(c)
```{r 2c}
mod2c <- lm(lwage ~ educ*black + exper + tenure + married + south + urban, data = wage2)
summary(mod2c)
```

Our null hypothesis is that the interaction term between education and black has no effect on the model response. 

$H0: \beta_8 =0$  
$H1: \beta_8 \neq 0$

Decision Rule: Reject H0 if |t| > 1.64   ($\alpha = 0.1 $, two sided test)

As t(educ:black) is |-1.121| < 1.64 (as well as: p-value of 26% > 10%), we cannot reject H0. The interaction term between education and black might not have an effect on the model. So, education might not depend on race.

#### 2(c) - i
```{r 2ci, results="asis"}
stargazer(mod2a,mod2c, title = "Comparison of models mod2a and mod2c", type="html", model.names = FALSE,column.labels = c("mod2a", "mod2c"), column.separate = c(1,1), style = "qje")
```

In model 2c we can see that the black coefficient is positive now. But this variable interacts with eduaction. This means being black causes an effect of $+0.09-0.023*educ$. So the wage of a black person depends in this case of his education. Model 2a do not integrate the interaction term. Therefore, being black decrease the persons wage by 19%.

$\frac{\delta log(wage)}{\delta black}=\beta_2+\beta_1educ$

However, the standard error for black in model 2c is quite big and we can't verify a significance for this coefficient anymore. As already seen in the hypothesis test above ($H0: \beta8 = 0$), education also does not depend on the variable black and the model 2a might be the one which reflects the population in a better way.

#### 2(c) - ii
```{r 2cii}
(c <- mean(wage2$educ))
educ_c <- wage2$educ - c

mod2cii <- lm(lwage ~ educ_c*black + exper + tenure + married + south + urban, data = wage2)
summary(mod2cii)
```

The value we suggest is equal to the mean of education (13.5). Interpreting models with interactions can be tricky due to the fact that $\beta2$ (coefficient of black) just shows the partical effect of being black when his eduaction = 0. While using the equation black*(educ-c) we centring the variables around the mean value of education. 

#### 2(d)
```{r 2d}
mod2d <- lm(lwage ~ educ + black*married + exper + tenure + south + urban, data = wage2)
summary(mod2d)
```

Being not married and not black is our base model, which is illustrated by the intercept. 
The second group (not_married & black) is represented by the coefficient $\beta2 = -0.240820$
The coefficient $\beta3 = 0.188915$ presents the third group: being married and not being black. 
Being married and black is represented by $\beta2 + \beta3 + \beta8$. 

Therefore, the difference between the second and the forth group are  $\beta2 + \beta8 = -0.24 + 0.06 = -0.18$. (Same as in mod2a)

##Task 3
```{r}
head(ksubs)
```

Just looking at the data "ksubs" reveals that our response variable e401k is a binary categorical variable. While fitting a linear regression model, it will lead to probabilities which are less than 0 or greater than 1. A logistic regression model fits better for this exercise.
-> Heteroskedastic???


####3(i)
```{r 3i}
(x <- table(ksubs$e401k))
n <- nrow(ksubs)
(Fractions <- x/n)
```
In our data we have 5638 people (61%) which are not eligible for participating in a 401(k) plan, and 3637 people (39%) which are eligible for participating. 

####3(ii): Linear Model

$e401k = \beta_0 + \beta_1*inc + \beta_2 * incsq +  \beta_3*age + \beta_4 * agesq + \beta_5 * male$
```{r 3ii}
mod3ii <- lm(e401k ~ inc + incsq + age + agesq + male, data = ksubs)
summary(mod3ii)
```


####3(iii + v) Interpretation of coefficients
The estimates in the summary output for each predictor are the effects on the response e401k if one predictor x is increased by one unit while holding the other predictors fixed (linear model).

$\frac{\delta e401k}{\delta inc}=\beta_1+\beta_2inc = 0.01245 -0.00006165 * inc$  
The effect of one unit increase in income depends on income itself. Saying we have an income about 30,000 we gain an effect of $-1.83705 = -183\%$.
&nbsp;
  
$\frac{\delta e401k}{\delta age}=\beta_3+\beta_4age = 0.02651 -0.0003053 * age$  
The effect for age is similar to variable income as the derivation also depends on the value itself. Saying we look at the age of 40 (median of age), the effect is $0.014298 = 1.4\%$. 
&nbsp;
  
$\frac{\delta e401k}{\delta male}=\beta_5 = -0.003533$
Our base model is being a woman. So being a male will increase the probability of eligibilty for participating by $-0.3533\%$.


####3(iv) Independence of e401k
```{r 3iv}
summary(mod3ii)
```
As shown in the output (mod3ii) we can assume that e401k is independent from gender. We test if $H_0: \beta_5 = 0$ in a two-sided test using $\alpha = 0.05$. As the p-value of 0.77 is bigger than 0.05 we cannot reject $H_0$. So, the probability of participating is independet of being male or female. 

Furthermore, the variable income and age are signficant for the model. Therefore, e401k is dependent on these both variables. We could also test these both hypothesis, which are based on the output of model 3ii:

Income:
$H_0: \beta_1 = 0, \beta_2 = 0$ (two sided test, $\alpha = 0.05$)
We reject $H_0$ as both p_values < 0.05. 

Age:
$H_0: \beta_3 = 0, \beta_4 = 0$ (two sided test, $\alpha = 0.05$)
We reject $H_0$ as both p_values < 0.05.

####3(v) Fitted Values
Because we are working with a categorical y variable and apply a linear model we are expecting to see values outside of the range 0 to 1.

```{r}
-5.063e-01 + 
mod3ii$coefficients["age"]
```

####3(vi) Confusion Matrix (1)
```{r 3 vi}
preds = predict(mod3ii)
head(preds)
# we would need a train and test set?? stupid prediction :D

pCutoff = 0.5
preds_cutoff = as.numeric(preds > pCutoff)

#table(preds_cutoff)
#table(ksubs$e401k)
(ConfMat = table(ksubs$e401k,preds_cutoff))
```
True Negatives = 4607 people are correctly classified not being eligible.
True Positives = 1429 people are correctly classified being eligible.

False Positives = 1031 non eligible people are classified as eligible. 
False Negatives = 2208 eligible people are classified as non eligible.

```{r}
(TP = sum(ConfMat[,2]))
```

In total, 2460 people are classfied as eligible.

####3(vii) Confusion Matrix (2)
```{r 3vii}
#Correctly predicted:
((4607+1429)/nrow(ksubs))
```


65% of all predicted observations are correctly classified.  

####3(viii)
```{r 3viii}
#Correctly predicted non_eligibles:
(TNR = ConfMat[1,1]/sum(ConfMat[,1]))

#Correctly predicted eligibles:
(TPR = ConfMat[2,2]/TP)
```
58% of all eligibiles are predicted correct.
68% of all non eligibles are predicted correct.

???
Welche Frage ist hier offen?